---
output: github_document
bibliography: vignettes/references.bib
csl:  vignettes/ieee.csl
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r, include = FALSE}
library(bibtex)
library(bookdown)

refs <- read.bib("vignettes/references.bib")
keys <- names(refs)
```

# allMT

<!-- badges: start -->

[![R-CMD-check](https://github.com/tmungle/allMT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tmungle/allMT/actions/workflows/R-CMD-check.yaml) [![Travis build status](https://travis-ci.com/tmungle/allMT.svg?branch=master)](https://travis-ci.com/tmungle/allMT) [![R-CMD-check](https://github.com/tmungle/allMT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tmungle/allMT/actions/workflows/R-CMD-check.yaml)
[![R-CMD-check](https://github.com/tmungle/allMT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tmungle/allMT/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Introduction

Acute lymphoblastic leukemia (ALL) studies across the globe have shown that delivering optimal maintenance therapy (MT) is crucial to achieve better treatment outcomes [@schmiegelow2014mercaptopurine]. Delivering optimal MT includes prescribing maximum tolerated doses of 6-mercaptopurine (6MP) and methotraxate (MTX) based on patient's blood counts (absolute neutrophil count (ANC), platelet count (PLT) and hemoglobin (HB)). In view of this, the *allMT* package was created to analyse and visualize the MT data (ANC/PLT/HB/6MP/MTX) for single patient or a given cohort at any center[^1]. The package was developed considering the ICiCLe-ALL-14 protocol [@das2022protocol] as default. It can be used to analyze data from centers with different MT protocols where in the user will need to alter the function defined ipnut argument/parameters.

[^1]: The work was initiated as part of Tushar Mungle's PhD work at IIT Kharagpur [@mungle2020modelling] and then continued at TTCRC, Tata Medical Center, Kolkata.

## Package

Using this package the user may:

1.  Convert maintenance sheets (Tata Medical Center Kolkata, India format or user dependent) into a cleaner, single csv with longitudinal data for each patient.

2.  Analyze MT [@mungle2020modelling]

    -   At an individual patient level

        -   To track patient's MT progression during or after therapy.

        -   To track patient's cycle-by-cycle MT progression during or after the therapy using summary measure (SM) weighted mean (6MP\*MTX) vs weighted mean ANC parameters in a scatterplot.

        -   Asses hematological toxicities - neutropenia, thrombocytopenia and anemia during MT.

        -   Evaluate median time to first 6MP dose increase for the cohort.

        -   Evaluate real time dosing decisions (stop, reduce or increase) by the physicians during MT based on MT dosing guidelines [@mungle2020modelling]

    -   At cohort level

        -   SM plot to analyze MT for a given cohort defined by the user. SM is evaluated for each patient and plotted together to represent the entire cohort.

        -   Compare SM for two or more cohort defined by the user to evaluate MT practice. Cohort comparisons can include clinical interventions, year-wise evolution, patients treated by two different set of physicians, or as user see it fit.

        -   Asses hematological toxicities - neutropenia, thrombocytopenia and anemia during MT.

        -   Plot depicting median time to first 6MP dose increase for the cohort.

        -   Evaluate real time dosing decisions (stop, reduce or increase) by the physicians during MT based on MT dosing guidelines [@mungle2020modelling]

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->

## Installation

You can install the development version of allMT from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tmungle/allMT")
```

## Ethics approval statement for data

The anonymized clinical data set refers to patients treated on the ICiCLe-ALL-14 treatment protocol at the Tata Medical Center Kolkata (institutional review board approval EC/TMC/12/13 for the treatment protocol). Funding support for the ICiCLe-ALL-14 clinical study (Clinical Trials Registry-India reference CTRI/2015/12/006434) was provided by the National Cancer Grid (2016/001; 2016-) and the Indian Council of Medical Research (79/159/2015/NCD-III; 2017-19)

## Contribution

By abiding to [Contributor Code of Conduct](https://www.contributor-covenant.org/version/2/1/code_of_conduct/), contribution to `allMT` in any form (bug fixes/code modification/new code development) is welcomed.

## References
