---
title: "Analyze Acute Lymphoblastic Leukemia Maintenance Therapy"
# subtitle: "Tushar Mungle$^1$$^,$$^2$, Ananya Mahadevan$^2$"
author:
  - name: "Tushar Mungle$^1$$^,$$^*$$^,$$^+$, Ananya Mahadevan$^1$$^,$$^*$, Shekhar Krishnan$^1$"
    affiliation: $^1$Tata Translational Cancer Research Centre (TTCRC), Tata Medical Center (TMC), Kolkata, India; $^*$contributed equally
    email: $^+$ Correspondence - tushar.mungle@gmail.com
    #email: $^*$Correspondence - tushar.mungle@gmail.com
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    number_sections: true
    toc_float: true
    # theme: cerulean
    code_folding: show
    code_download: TRUE
vignette: >
  %\VignetteIndexEntry{Analyze Acute Lymphoblastic Leukemia Maintenance Therapy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
csl: ieee.csl
link-citations: true
# css: "../styles.css"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(bibtex)
library(bookdown)

refs <- read.bib("references.bib")
keys <- names(refs)
```

------------------------------------------------------------------------

# Introduction

***Acute Lymphoblastic Leukemia*** (ALL)

ALL treatment is broadly divided into 2 phases - Intensive Phase and Maintenance Phase. The intensive phase of treatment is divided into 4 sub-phases -- induction, consolidation, interim maintenance and delayed intensification [@das2022protocol].

Maintenance therapy (MT), the longest phase of ALL treatment consists of \~2 years (96 weeks or 8 cycles - 12 weeks per cycle) [@das2022protocol]. Patients are instructed to visit the hospital once every two weeks for consultation and on a weekly basis in the event of toxicities (neutropenia, thrombocytopenia or any other). Overall, a patient should ideally have a minimum of 48 visits during MT. During each visit, patient's blood report is examined for the following parameters - absolute neutrophil count (ANC), platelet count (PLT) and hemoglobin (HB) to adjust the drugs 6-mercaptopurine (6MP) and Methotrexate (MTX) that are prescribed on weekly basis.

------------------------------------------------------------------------

# Package

Studies across the globe have shown that delivering optimal MT is crucial to achieve better treatment outcomes [@schmiegelow2014mercaptopurine]. In view of this, the package was created to analyse and visualize the MT data (ANC/PLT/HB/6MP/MTX) of ALL therapy at any given center[^1]. Although, the package was developed considering the ICiCLe-ALL-14 protocol [@das2022protocol] as default, it can be used to analyze data from centers with different MT protocols. In case of different protocol, the user will need to alter the defined parameters specific to protocol as function arguments.

[^1]: The work was initiated as part of Tushar Mungle's PhD at IIT Kharagpur [@mungle2020modelling] and then continued at TTCRC, Tata Medical Center, Kolkata.

Using this package the user may:

1.  Convert maintenance sheets (Tata Medical Center Kolkata, India format or user dependent) into a cleaner, single csv with longitudinal data for each patient.

2.  Analyze MT

    -   At an individual patient level

        -   To track patient's MT progression during or after therapy [@mungle2020modelling].

        -   To track patient's cycle-by-cycle MT progression during or after the therapy using summary measure (SM) weighted mean (6MP\*MTX) vs weighted mean ANC parameters in a scatterplot.

    -   At cohort level

        -   SM plot to analyze MT for a given cohort defined by the user. SM is evaluated for each patient and plotted together to represent the entire cohort [@mungle2020modelling].

        -   Compare SM for two different cohort defined by the user to evaluate MT practice. Cohort comparisons can include clinical interventions, year-wise evolution, patients treated by two different set of physicians, or as user see it fit [@mungle2020modelling].

3.  Evaluate real time dosing decisions (stop, reduce or increase) by the physicians during MT based on MT dosing guidelines for the individual patient or cohort [@mungle2020modelling].

4.  Evaluate hematological toxicities - neutropenia, thrombocytopenia and anaemia during MT for the individual patient or cohort.

5.  Evaluate/plot median time to first 6MP dose increase for the individual patient or cohort [@mungle2020modelling].

```{r setup}
# library(allMT)
```

------------------------------------------------------------------------

# Data

1.  The data recording sheet used by TMC can be downloaded/copied as

    ```{r}
    # Uncomment below line before running the snippet
    # library(allMT)
    pat_data <- system.file("extdata/tmc_data/", "UPN_916.xls", package = "allMT")
    dest_path <- getwd() # user may choose a different destination 
    file.copy(pat_data, dest_path)
    ```

2.  To use functions of the package, the data needs to be in a specific format as described below. Sample data[^2] for 10 individual patient is available as UPN914, UPN915,...,UPN923 under the folder "/extdata/processed_data/".

    Data can be checked or saved as

    ```{r}
    # Working example (read.csv() does not recognize the path via system.file() in rmd, hence absolute path)
    pat_df <- read.csv("C:/Users/Shubham/Desktop/UPN_915.csv", header = TRUE, sep = ",")
    head(pat_df)

    # Below code should work for any user independent of OS
    #pat_data <- system.file("extdata/processed_data", "UPN_915.csv", package = "allMT")
    #pat_df <-read.csv(pat_data, header = TRUE, sep = ",")
    #head(pat_df)


    ```

[^2]: Ethics approval statement: The pseudonymised clinical dataset used in this library included patients treated at the Tata Medical Center Kolkata on the ICiCLe-ALL-14 clinical trial (Clinical Trials Registry-India CTRI/2015/12/006434). Funding support for the clinical trial was provided by the National Cancer Grid (2016/001; 2016-) and the Indian Council of Medical Research (79/159/2015/NCD-III; 2017-19).

------------------------------------------------------------------------

# Functions

## Convert Data

-   `convert_tmc_format()` - Convert information recorded in the TMC datasheet format to a format that will be used by package functions.

-   `convert_external_format()` - Convert MT information (user defined) to a format that will be used by package functions.\
    [*Note:*]{.underline} User needs to ensure the MT information is stored in an excel or csv with column names as - Dates, ANC, PLT, MP, MTX. The function would convert this sheet into required format.

The output of both the functions will result in standard patient csv files (as shown in Data section - head(pat_df()) saved in a folder defined by the user.

## Analyze MT Data

### At individual patient level

-   `plot_progression()` - Plot line graph showing historical longitudinal trends of ANC, 6MP and MTX for a single patient. This function can be use by the doctors during the MT clinic to monitor patients' MT dose prescription and treatment response in real time.

-   `summarize_cycle_progression()` - Plot SM data per cycle for given patient data. Utility as above.

### At cohort level

-   `summarize_cohortMT()` - Plot scatter plot (SM) for all the patients from user defined cohort.

-   `compare_cohorts()` - Plot scatter plot (SM) for all the patients present in user defined cohort 1 against cohort 2. Two methods of cohort comparison can be considered for analysis as mentioned below:

    -   Method 1 - Comparison between cohorts that began MT before and after a particular user-provided date.

    -   Method 2 - Comparison between 2 cohorts pre-determined by the user.

## Evaluate Physicians' Dosing Decisions

-   `assess_stop_doses()` - The function evaluates the incidence of 'No Dose' physician's decisions that are NOT supported by protocol-based target blood counts.

-   `assess_reduced_doses()` - The function evaluates the incidence of 'Reduce Dose' physician's decisions that are NOT supported by protocol-based target blood counts.

-   `assess_increased_doses()` - The function evaluates - The function evaluates the incidence of 'Increase Dose' physician's decisions that are NOT supported by protocol-based target blood counts.

    [*Note:*]{.underline} Depending upon value provided for the argument 'inputfolder_path' as a

    1.  path to a file for single patient csv - resulting output will be analysis for single patient
    2.  path to a folder (cohort) with multiple csv - resulting output will analysis carried for the cohort of patients' csv files present in the folder

## Evaluate Hematological Toxicities

-   `assess_neutropenia()` - The function evaluates number of neutropenia episodes as per the neutropenia definition provided in 'anc_range' argument - c(a,b). a is defined as highest value of ANC threshold for neutropenic condition, and b is defined as ANC value above which neutropenic condition is said to be recovered.

-   `assess_thrombocytopenia()` - The function evaluates number of thrombocytopenia episodes as per the thrombocytopenia definition provided in 'plt_range' argument - c(a,b). a is defined as highest value of PLT threshold for thrombocytopenia condition, and b is defined as PLT value above which thrombocytopenic condition is said to be recovered.

`assess_anaemia()` - The function evaluates number of anaemia episodes as per the anaemia definition provided in 'hb_range' argument - c(a,b). a is defined as highest value of Hb threshold for anaemia condition, and b is defined as Hb value above which anaemic condition is said to be recovered.

[*Note:*]{.underline} Depending upon value provided for the argument 'inputfolder_path' as a

1.  path to a file for single patient csv - resulting output will be analysis for single patient
2.  path to a folder (cohort) with multiple csv - resulting output will analysis carried for the cohort of patients' csv files present in the folder

## Time to First-Dose-Increase

-   `time_to_doseinc()` - Plot the graph showing median time to first 6-MP dose increase for the cohort (if input file is an individual patient instead of cohort, function displays value for an individual patient in question)

------------------------------------------------------------------------

# References
